Handling missing data in smartphone location logs

### 2017-11-17 thesisReport.Rmd
added a bit of meat to the methods section

The data used was collected between 2014 and 2017 on different Android devices from X individuals across X continents. The table below provides more details:

```{r datadetailsTable, echo = F, fig.cap="caption here?"}
datadetails <- readRDS("tempdata/datadescriptives.rds")
datadetails <- datadetails[,-7]
kable(datadetails,digits = 2,col.names = c("Log duration",
                                           "Days of log",
                                           "Observations",
                                           "Missing days",
                                           "Missing data",
                                           "Mean Accuracy"))
```

In addition to the secondary logs, participants also volunteered to carry with them a specialised GPS tracker for week. Therefore we also used a week of parallel secondary and specialised logs.

@@

The data was discretised into 5 minute windows following @palmius_detecting_2017 (as described above, for more detail please consult the original paper).

@@

Four imputation methods were selected in order to cover techniques applied in the literature. Briefly, the filtered mean imputation method described by @palmius_detecting_2017, the model developed by @barnett_inferring_2016 as well as an adapted FarOut model [@sadilek_far_2016]. Simple Kalman filtered linear interpolation was used as a benchmark models.

@@

Measures of performance used were root mean square error (RMSE) and mean absolute error (MAE).

@@

![Raw GPS measurements of a journey from de Uithof to Tuinwijk on February 17th 2017. The measurements are in red, the filtered path is in blue. The circles denote 67% confidence intervals of the given GPS measurement. Measurements and fitted points which follow each other in time are connected by lines. The inacurate measurements lead to estimates of irregular movements. The filtered movement estimate seems more accurate in terms of the path, but lags behind the unfiltered movement.](img/Map1.png)

>>>>

The data used to train the imputation methods was collected between 2013 and 2017 on different Android devices from several individuals. The table below provides more details:

```{r datadetailsTable, echo = F}
datadetails <- readRDS("tempdata/datadescriptives.rds")
datadetails <- datadetails[,-7] #remove last column for now
kable(datadetails,digits = 2,col.names = c("Log duration",
                                           "Logged days",
                                           "Observations",
                                           "Missing days",
                                           "Missing data",
                                           "Mean Accuracy"),
      caption = "Table with descriptives about the data sets used to build the imputation methods. ")
```

In addition to the secondary logs, participants also volunteered to carry with them a specialised GPS tracker for a week. This specialised log was used to evaluate the models.

@@

Three different filtering methods were tested. First, the filtered rolling-median downsampling method described by @palmius_detecting_2017. The second method is a weighted mean approach taking $f(a)$ as a weight. Finally, an extended Kalman filter. The output of all of these methods was taken as the input of the imputation methods.

@@

Four imputation methods were selected in order to cover techniques applied in the literature. Briefly, the mean imputation method described by @palmius_detecting_2017, the model developed by @barnett_inferring_2016 as well as an adapted FarOut model [@sadilek_far_2016]. Simple linear interpolation was used as a benchmark model.

@@

The direct evaluation involved calculating the error of each $D_t$ compared to $G(t)$ approximated by the specialised log. The error measures used were root mean square error (RMSE) and mean absolute error (MAE).

The evaluation on computed measures involved calculating a mobility trace following the rectangual method of @rhee_human_2007 for each imputed dataset. Like @barnett_inferring_2016 I calculate bias by substracting the estimated measure under each approach for the same measure calculated on the full data. For simulation-based imputation approaches a mean value over 100 samples was taken.

@@
